Improved Approximation Algorithms for the Maximum Carpool Matching Problme

### 2017-02-15 src/abstract.tex
merge dror

We show that \carpool can be formulated as an unconstrained submodular
maximization problem, thus it admits a $\frac{1}{2}$-approximation
algorithm.  We show that the same formulation does not work for \gcp,
nevertheless, we present a $(\frac{1}{2} - \eps)$-approximation
algorithm for \gcp.
%
For the unweighted variant of both problems when the maximum possible
capacity, $\cmax \eqdf \max_{v \in V} c(v)$, is bounded by a constant,
we provide a $(\half + \inv{2\cmax} - \eps)$-approximation algorithm.
We also show that the problem remains APX-hard, even if $\cmax \leq
3$.

>>>>

We show that \carpool can be formulated as an unconstrained submodular
maximization problem, thus it admits a $\frac{1}{2}$-approximation
algorithm.  We show that the same formulation does not work for \gcp,
nevertheless, we present a local search $(\frac{1}{2}
- \eps)$-approximation algorithm for \gcp.
%
For the unweighted variant of both problems when the maximum possible
capacity, $\cmax$, is bounded by a constant, we provide a local search
$(\half + \inv{2\cmax} - \eps)$-approximation algorithm.  We also show
that the problem APX-hard, even if $\cmax \leq 3$.

### 2017-02-15 src/appendix.tex
merge dror

\section{Omitted Proof}
\begin{proof}[Proof of Theorem~\ref{thm:improve}]
Define the weight function $w'(e)
= \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps}}$, where
$w_{\max} \eqdf \max_{e \in A} w(v)$.
%
Let $M^*$ and $M'$ be an optimal carpool matchings with respect to $w$
and $w'$, resp.  We have that
\[
w'(M^*)
=    \sum_{e \in M^*} \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps}}
>    \sum_{e \in M^*} \paren{ \frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps} } - m
=    \frac{m}{2\eps w_{\max}} \cdot w(M^*) - m
~.
\]
Since $w(M^*) \geq w_{\max}$, we have that 
\[
w'(M^*)
>    \frac{m}{2\eps w_{\max}} \cdot (w(M^*) - 2\eps w_{\max})
\geq \frac{m}{2\eps w_{\max}} \cdot (1 - 2\eps) w(M^*)
~.
\]
By Theorem~\ref{thm:improve-bounded}, our local improvement algorithm computes a
$\half$-approximate carpool matching $M$ on $(G, c, w')$, and this
matching satisfies $w'(M) \geq w'(M')/2$.  Furthermore, since
$w'(M') \geq w'(M^*)$, it follows that $w'(M) \geq w'(M^*)/2$.
Therefore
\[
w(M)
\geq \frac{2\eps w_{\max}}{m} w'(M) 
\geq \frac{1}{2} \frac{2\eps w_{\max}}{m} w'(M^*)
>    \frac{1-2\eps}{2} w(M^*)
~,
\]
as required.
\end{proof}

\section{Tight Analysis}
We now show that Corollary~\ref{cor:local} is tight.  Consider the
example in Figure~\ref{fig:local search tight}, the example is for the
special case when $k = 11$ and $\cmax = 2$ but this example can be
generalized in a straightforward manner.  One can verify that as the
graph in the example growth, the ratio between the optimal solution
and the local search solution approaches $\frac{3}{4}$.

\begin{figure}
\begin{center}
\scalebox{.85}{\input{fig-local-tight}}
\caption{A tight example for Corollary~\ref{cor:local}. 
The optimal solution is given by the red solid arcs while the local
search solution is given by the green, dashed arcs.  The solution can
be improved by the local search algorithm only if it removes all the
edges.}
\label{fig:local search tight}
\end{center}
\end{figure}

>>>>

\section{Omitted Proofs}
\label{sec:omitted}

\begin{proof}[Proof of Theorem~\ref{thm:improve}]
Define the weight function $w'(e)
= \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps}}$, where
$w_{\max} \eqdf \max_{e \in A} w(v)$.
%
Let $M^*$ and $M'$ be an optimal carpool matchings with respect to $w$
and $w'$, resp.  We have that
\[
w'(M^*)
=    \sum_{e \in M^*} \floor{\frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps}}
>    \sum_{e \in M^*} \paren{ \frac{w(e)}{w_{\max}} \cdot \frac{m}{2\eps} } - m
=    \frac{m}{2\eps w_{\max}} \cdot w(M^*) - m
~.
\]
Since $w(M^*) \geq w_{\max}$, we have that 
\[
w'(M^*)
>    \frac{m}{2\eps w_{\max}} \cdot (w(M^*) - 2\eps w_{\max})
\geq \frac{m}{2\eps w_{\max}} \cdot (1 - 2\eps) w(M^*)
~.
\]
By Theorem~\ref{thm:improve-bounded}, our local improvement algorithm computes a
$\half$-approximate carpool matching $M$ on $(G, c, w')$, and this
matching satisfies $w'(M) \geq w'(M')/2$.  Furthermore, since
$w'(M') \geq w'(M^*)$, it follows that $w'(M) \geq w'(M^*)/2$.
Therefore
\[
w(M)
\geq \frac{2\eps w_{\max}}{m} w'(M) 
\geq \frac{1}{2} \frac{2\eps w_{\max}}{m} w'(M^*)
>    \frac{1-2\eps}{2} w(M^*)
~,
\]
as required.
\end{proof}

\medskip

\begin{proof}[Proof of Lemma~\ref{lemma:dec}]
We give a constructive proof.
%
First, if $\ell \leq |V| \leq \Delta \ell$, then we are done.
%
Otherwise, we partition the graph recursively as follows.  Let $U
= \emptyset$.  As long as $\abs{U} < \ell$, choose an arbitrary vertex
$u$ such that $U \cup \set{u}$ is connected and add $u$ to $U$.
%
If the graph $G[V \setminus U]$, which is induced by $V \setminus U$,
is connected, then choose the next vertex.
%
However, if $G[V \setminus U]$ becomes disconnected, then consider the
maximal connected components of $G[V \setminus U]$.  Add the vertices
of any maximal component that contains less than $\ell$ vertices to
$U$.  Observe that afterwards $\abs{U} < \ell + (\Delta-1) \ell
= \Delta \ell$.
%
If $\abs{U} \geq \ell$, then recursively partition any component of
$G[V \setminus U]$ that contains more than $\ell$ vertices.
%
If $\abs{U} < \ell$, then there must exist at least one maximal
component $C$ of $G[V \setminus U]$ that contains more than $\ell$
vertices.  In this case recursively partition any maximal component of
$G[V \setminus (U \cup C)]$ that contains more than $\ell$ vertices.
Also, continue to augment $U$ in the graph $G[U \cup C]$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Figures}
\label{sec:figures}

\begin{figure}[h]
\centering
\input{fig-submodular}
\caption[]{%
We show that the arcs from \subref{sub:cup} and \subref{sub:cap} can be
redistributed over \subref{sub:s} and \subref{sub:s} 
%\todo{write explanation}
}
\label{fig:sub}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section{Tight Analysis}
\label{sec:tight}

We now show that Corollary~\ref{cor:local} is tight.  Consider the
example in Figure~\ref{fig:local search tight}, the example is for the
special case when $k = 11$ and $\cmax = 2$ but this example can be
generalized in a straightforward manner.  One can verify that as the
graph in the example growth, the ratio between the optimal solution
and the local search solution approaches $\frac{3}{4}$.

\begin{figure}[h]
\begin{center}
\scalebox{.85}{\input{fig-local-tight}}
\caption{A tight example for Corollary~\ref{cor:local}. 
The optimal solution is given by the red solid arcs while the local
search solution is given by the green, dashed arcs.  The solution can
be improved by the local search algorithm only if it removes all the
edges.}
\label{fig:local search tight}
\end{center}
\end{figure}

### 2017-02-15 src/group-apx.tex
merge dror

We now argue that the local improvement algorithm from
Section~\ref{sec:improve} can be generalized to solve \gcp.
%
The main concern when trying to adopt the algorithm to \gcp is how to
determine if a vertex can be improved.
%
With \carpool, if weights are polynomially-bounded, it was enough to
consider the incoming arcs to a vertex $v$ in a non-increasing
order of $\delta_M$ (See proof of Theorem~\ref{thm:improve-bounded}).
This does not work anymore, since in the \gcp we have sizes.  In fact, given
$v$, finding the best star with respect to $\delta_M$ is
a \textsc{Knapsack} instance where the size of the knapsack is $c(v)$.
%
If weights are polynomially-bounded, then $\delta_M(e)$ is bounded for
every arc $e \in A$, and therefore this instance of \textsc{Knapsack}
can be solved in polynomial time using dynamic programming.

\begin{theorem}
Algorithm~\ref{alg:grd} is a $\half$-approximation algorithm for \gcp,
if edge weights are integral and polynomially bounded.
\end{theorem}

>>>>

We now argue that Algorithm~\textbf{StarImprove} from
Section~\ref{sec:improve} can be generalized to solve \gcp.
%
The main concern when trying to adopt the algorithm to \gcp is how to
determine if a vertex can be improved.
%
With \carpool, if weights are polynomially-bounded, it was enough to
consider the incoming arcs to a vertex $v$ in a non-increasing
order of $\delta_M$ (See proof of Theorem~\ref{thm:improve-bounded}).
This does not work anymore, since in the \gcp we have sizes.  In fact, given
$v$, finding the best star with respect to $\delta_M$ is
a \textsc{Knapsack} instance where the size of the knapsack is $c(v)$.
%
If weights are polynomially-bounded, then $\delta_M(e)$ is bounded for
every arc $e \in A$, and therefore this instance of \textsc{Knapsack}
can be solved in polynomial time using dynamic programming.

\begin{theorem}
Algorithm~\textbf{StarImprove} is a $\half$-approximation algorithm
for \gcp, if edge weights are integral and polynomially bounded.
\end{theorem}

### 2017-02-15 src/hardness.tex
done my pass
fix
merge dror

??????

As we mentioned in the introduction, the maximum spanning star forest
can not be approximated within a factor of $\frac{31}{32} + \epsilon$
for any $\epsilon > 0$, unless P$=$NP.  The proof, however, does not
hold for the case when the maximum capacity is constant.  In this
subsection we show that the problem remains APX-hard even in this
case.

??????

@@

\begin{theorem}
The \carpool problem is APX-hard, even for undirected and unweighted
instances with maximum capacity bounded by $b$, for $b \geq 3$.
\end{theorem}
\begin{proof}
We prove the theorem by presenting an $L$-reduction from
\textsc{Minimum Dominating Set-$b$}.
(For details on $L$-reductions the reader is referred to~\cite{??}.)

@@

Let $D^*$ be an optimal dominating set with respect to $G$, and let
$M^*$ be an optimal carpool matching with respect to $G$ and $c$.
Since $|D^*| \geq \frac{|V|}{b+1}$, it follows that 
\[
\abs{P_{M^*}} \leq b \abs{D^*}
~,
\]
In addition, if $M$ is a carpool matching, we have that
\[
\abs{D_M \cup Z_M} - \abs{D*}
= (\abs{V} - \abs{P_M}) - \abs{D^*}
= \abs{P_{M^*}} - \abs{P_M}
~.
\]
Hence, there is an $L$-reduction from \textsc{Minimum Dominating
Set-$b$} to unweighted and undirected \carpool with bounded capacity
$b$.
\end{proof}

>>>>

As we mentioned earlier, \textsc{Spanning Star Forest} has a lower
bound of $\frac{10}{11} + \eps$ for any $\eps > 0$, unless
P$=$NP~\cite{ChakrabartyGoel10}, and this bound applies to \carpool.
The result, however, does not hold for the case where $\cmax = O(1)$.
In this section we show that the problem remains APX-hard even in this
case.

@@

\begin{theorem}
The \carpool problem is APX-hard, even for undirected and unweighted
instances with maximum capacity bounded by $b$, for $b \geq 3$.
\end{theorem}
\begin{proof}
We prove the theorem by presenting an $L$-reduction from
\textsc{Minimum Dominating Set-$b$}.
(For details on $L$-reductions the reader is referred to~\cite{PapYan88}.)

@@

Let $D^*$ be an optimal dominating set with respect to $G$, and let
$M^*$ be an optimal carpool matching with respect to $G$ and $c$.
Since $|D^*| \geq \frac{|V|}{b+1}$, it follows that 
\(
\abs{P_{M^*}} \leq b \abs{D^*}
%~,
\),
In addition, if $M$ is a carpool matching, we have that
\(
\abs{D_M \cup Z_M} - \abs{D^*}
= (\abs{V} - \abs{P_M}) - \abs{D^*}
= \abs{P_{M^*}} - \abs{P_M}
%~.
\).
Hence, there is an $L$-reduction from \textsc{Minimum Dominating
Set-$b$} to unweighted and undirected \carpool with bounded capacity
$b$.
\end{proof}

### 2017-02-15 src/improve.tex
merge dror

In this subsection we give a local improvement
$(\half-\eps)$-approximation algorithm for \carpool.  This algorithm
repeatedly checks whether the current carpool matching $M$ can be
improved by means of a star centered at a vertex $v$.  The profit from
this star is the total weight of the arcs in the star, and the cost is
the total weight of lost arcs (e.g., arcs from passengers to drivers
that became passengers of $v$).  If the profit is larger than the
cost, then an improvement step is performed.  The algorithm terminates
when such a step is not possible.
%
We remind the reader that this algorithm will be extended to \gcp in
Section~\ref{sec:group}.

We need a few definitions before presenting our algorithm.
%
Given a directed graphs $G = (V,A)$, define $\nin[] \eqdf \set{u :
(u,v) \in A}$ and $\nout[] \eqdf \set{u : (v,u) \in A}$.
%
Let $M$ be a feasible carpool matching.  The weight $w_M(v)$ of a
vertex $v$ with respect to $M$ is the sum of the weights of the arcs
in $M$ that are incident on $v$, namely
\[
\textstyle
w_M(v)
\eqdf w(M \cap \nin[]) + w(M \cap \nout[])
=     \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u)
~.
\]
For a subset of vertices $U \subseteq V$ we define
$w_M(U) \eqdf \sum_{v \in U} w_M(v)$.

@@

\begin{observation}
\label{lm:val-twice}
$w_M(V) = 2 w(M)$.
%\sum_{v \in V} w_M(v) = 2 \sum_{e \in M} w(e)$.
\end{observation}
\begin{proof}
\(
\sum_{v \in V} w_M(v)
= \sum_{v \in V} \sum_{(u, v) \in M} w(u, v) +
    \sum_{v \in V} \sum_{(v, u) \in M} w(v, u) 
= 2 \sum_{e \in M} w(e)
\).
\end{proof}

@@

\begin{definition}
A subset $S_v \subseteq A \cap (V \times \{v\})$ is
an \emph{improvement} with respect to a carpool matching $M$, if the
following conditions hold:%
\begin{inparaenum}[(i)]
\item $\abs{S_v} \leq c(v)$, and
\item $\delta_M(S_v) > w_M(v)$.
\end{inparaenum}
Furthermore, if there exists an improvement for a vertex $v$, we say
that vertex $v$ can be \emph{improved}.
\end{definition}

@@

We are now ready to describe our local improvement algorithm
(Algorithm~\ref{alg:grd}).  The algorithm starts with an empty carpool
matching $M$, and in every iteration it looks for a vertex that can be
improved.  If there exists such a vertex $v$, then the algorithm
removes the arcs that are incident on it from $M$, and adds the arcs
in $S_v$.  The algorithm terminates when no vertex can be improved.
Figure~\ref{fig:improvement} depicts an improvement step.

\begin{algorithm}
\caption{Local Improvement}
\label{alg:grd}
$M \gets \emptyset$ \\
\Repeat{$\text{done}$}{
	$\text{done} \leftarrow \textsc{True}$ \\
	\For{$v \in V$}{
		\If{there exists an improvement $S_v$}{
			$M \leftarrow M \setminus \inc(S_v) \cup S_v$	\\
			$\text{done} \leftarrow \textsc{False}$ \\
		}
	}
}
\end{algorithm}

@@

\begin{lemma}
\label{lm:no improve}
Let $M$ be a matching computed by the local improvement algorithm.
Let $v$ be a vertex with no improvement, and let
$S \subseteq \nin(v)$, such that $\abs{S} \leq c(v)$, then $w(S) \leq
w_M(v) + w_M(\nin[S](v))$.
\end{lemma}
\begin{proof}
If no improvement exists, then we have that \\
\(
w(S) - w_M(\nin[S](v))
=    \sum_{(u,v) \in S} (w(u,v) - w_M(u))
=    \delta_M(S) 
\leq w_M(v)
~.
\)
\end{proof}

@@

\begin{lemma}
If Algorithm~\ref{alg:grd} terminates, then the computed solution is a
$\half$-approximation.
\end{lemma}
\begin{proof}
Let $M$ be the matching produced by the algorithm, and let $M^*$ be an
optimal matching.  We load every vertex $v$ with an amount of money
equal to $w_M(v)$, and then we show that this is enough to pay for
every arc in the optimal matching.  Due to
Observation~\ref{lm:val-twice} the total amount of money that we use
is exactly twice the weight of $M$.

@@

\begin{figure}
\centering
\input{fig-2-tight}
\caption[]{%
Consider a path with $2n + 1$ arcs, and alternating arc weights (2 and
1), if the local improvement algorithm selects all arcs of weight 1,
then no further improvement can be done and the value of the matching
is $n + 1$, while the optimal matching has value of $2n$.}
\label{fig:grd worst}
\end{figure}

@@

\begin{theorem}
\label{thm:improve-bounded}
Algorithm~\ref{alg:grd} is a $\half$-approximation algorithm
for \carpool, if edge weights are integral and polynomially bounded.
\end{theorem}
\begin{proof}
First, observe that determining if a vertex $v$ can be improved can be
done efficiently by considering the incoming arcs to $v$ in a
non-increasing order of their $\delta_M$s, and only ones with positive
values.  A vertex $v$ can be improved, then, if the $\delta$s of the
first $c(v)$ (or less) arcs sum up to more than $w_M(v)$.
%
It follows that the running time of an iteration of the for-loop is
polynomial.  Since the edge weights are integral and polynomially
bounded, the weight of an optimal carpool matching is polynomially
bounded.  The algorithm runs in polynomial time, because in each
iteration the algorithm improves the weight of the matching by at
least one or otherwise it terminates.
\end{proof}

It remains to consider the case, where weight are not polynomially
bounded.  In this case one can use scaling and rounding to ensure a
polynomial running time in the cost of a $(1+\eps)$ factor in the
approximation ratio.
%
The proof of the next theorem is given in the appendix.

>>>>

In this subsection we give a local search $(\half-\eps)$-approximation
algorithm for \carpool.  This algorithm repeatedly checks whether the
current carpool matching $M$ can be improved by means of a star
centered at a vertex $v$.  The profit from this star is the total
weight of the arcs in the star, and the cost is the total weight of
lost arcs (e.g., arcs from passengers to drivers that became
passengers of $v$).  If the profit is larger than the cost, then an
improvement step is performed.  The algorithm terminates when such a
step is not possible.
%
We remind the reader that this algorithm will be extended to \gcp in
Section~\ref{sec:group}.

We need a few definitions before presenting our algorithm.
%
Given a directed graphs $G = (V,A)$, define $\nin[] \eqdf \set{u :
(u,v) \in A}$ and $\nout[] \eqdf \set{u : (v,u) \in A}$.
%
Let $M$ be a feasible carpool matching.  The weight $w_M(v)$ of a
vertex $v$ with respect to $M$ is the sum of the weights of the arcs
in $M$ that are incident on $v$, namely
\[
%\textstyle
w_M(v)
\eqdf w(M \cap \nin[]) + w(M \cap \nout[])
=     \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u)
~.
\]
For a subset of vertices $U \subseteq V$ we define
$w_M(U) \eqdf \sum_{v \in U} w_M(v)$.

@@

\begin{observation}
\label{lm:val-twice}
$w_M(V) = 2 w(M)$.
%\sum_{v \in V} w_M(v) = 2 \sum_{e \in M} w(e)$.
\end{observation}
\begin{proof}
\(
\displaystyle
\sum_{v \in V} w_M(v)
= \sum_{v \in V} \sum_{(u, v) \in M} w(u, v) +
    \sum_{v \in V} \sum_{(v, u) \in M} w(v, u) 
= 2 \sum_{e \in M} w(e)
\).
\end{proof}

@@

\begin{definition}
A subset $S_v \subseteq A \cap (V \times \{v\})$ is
an \emph{improvement} with respect to a carpool matching $M$, if%
%the following conditions hold:%
\begin{inparaenum}[(i)]
\item $\abs{S_v} \leq c(v)$, and
\item $\delta_M(S_v) > w_M(v)$.
\end{inparaenum}
Furthermore, if there exists an improvement for a vertex $v$, we say
that vertex $v$ can be \emph{improved}.
\end{definition}

@@

We are now ready to describe our local search algorithm, which is
called \textbf{StarImprove} (Algorithm~\ref{alg:grd}).  It starts with
an empty carpool matching $M$, and in every iteration it looks for a
vertex that can be improved.  If there exists such a vertex $v$, then
the algorithm removes the arcs that are incident on it from $M$, and
adds the arcs in $S_v$.  The algorithm terminates when no vertex can
be improved.  Figure~\ref{fig:improvement} depicts an improvement
step.

\begin{algorithm}
\caption{\textbf{StarImprove}$(G,c)$}
\label{alg:grd}
\begin{small}
$M \gets \emptyset$ \\
\Repeat{$\text{done}$}{
	$\text{done} \leftarrow \textsc{True}$ \\
	\For{$v \in V$}{
		\If{there exists an improvement $S_v$}{
			$M \leftarrow M \setminus \inc(S_v) \cup S_v$	\\
			$\text{done} \leftarrow \textsc{False}$ \\
		}
	}
}
\end{small}
\end{algorithm}

@@

\begin{lemma}
\label{lm:no improve}
Let $M$ be a matching computed by \textbf{StarImprove}.  Let $v$ be a
vertex with no improvement, and let $S \subseteq \nin(v)$, such that
$\abs{S} \leq c(v)$, then $w(S) \leq w_M(v) + w_M(\nin[S](v))$.
\end{lemma}
\begin{proof}
If no improvement exists, then we have that \\
\(
w(S) - w_M(\nin[S](v))
=    \sum_{(u,v) \in S} (w(u,v) - w_M(u))
=    \delta_M(S) 
\leq w_M(v)
~.
\)
\end{proof}

@@

\begin{lemma}
If \textbf{StarImprove} terminates, then the computed solution is
$\half$-approximate.
\end{lemma}
\begin{proof}
Let $M$ be the matching produced by the algorithm, and let $M^*$ be an
optimal matching.  We load every vertex $v$ with an amount of money
equal to $w_M(v)$, and then we show that this is enough to pay for
every arc in the optimal matching.  Due to
Observation~\ref{lm:val-twice} the total amount of money that we use
is exactly twice the weight of $M$.

@@

\begin{figure}
\centering
\input{fig-2-tight}
\caption[]{%
Consider a path with $2n + 1$ arcs, and alternating arc weights (2 and
1), if \textbf{StarImprove} selects all arcs of weight 1, then no
further improvement can be done and the value of the matching is $n +
1$, while the optimal matching has value of $2n$.}
\label{fig:grd worst}
\end{figure}

@@

\begin{theorem}
\label{thm:improve-bounded}
Algorithm~\textbf{StarImprove} is a $\half$-approximation algorithm
for \carpool, if edge weights are integral and polynomially bounded.
\end{theorem}
\begin{proof}
First, observe that determining if a vertex $v$ can be improved can be
done efficiently by considering the incoming arcs to $v$ in a
non-increasing order of their $\delta_M$s, and only ones with positive
values.  A vertex $v$ can be improved, then, if the $\delta$s of the
first $c(v)$ (or less) arcs sum up to more than $w_M(v)$.
%
It follows that the running time of an iteration of the for-loop is
polynomial.  Since the edge weights are integral and polynomially
bounded, the weight of an optimal carpool matching is polynomially
bounded.  The algorithm runs in polynomial time, because in each
iteration the algorithm improves the weight of the matching by at
least one or otherwise it terminates.
\end{proof}

It remains to consider the case, where weight are not polynomially
bounded.  In Appendix~\ref{sec:omitted} we prove that one can use
standard scaling and rounding to ensure a polynomial running time in
the cost of a $(1+\eps)$ factor in the approximation ratio.

### 2017-02-15 src/in/title.tex
refactor
merge dror

\title{Improved Approximation Algorithms for the
Maximum Carpool Matching Problem}

\titlerunning{Improved Approximation Algorithms for Maximum Carpool Matching}

@@

>>>>

\title{%Improved Approximation
Local Search Algorithms for the Maximum Carpool Matching Problem}

\titlerunning{%Improved
Local Search Approximation Algorithms for Maximum Carpool Matching}

@@

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\EventEditors{John Q. Open and Joan R. Acces}
%\EventNoEds{2}
%\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
%\EventShortTitle{CVIT 2016}
%\EventAcronym{CVIT}
\EventYear{2017}
%\EventDate{December 24--27, 2016}
%\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{1} %42
\ArticleNo{1} %23
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-02-15 src/intro.tex
merge dror

\begin{figure}
\centering
\input{fig-carpool}
\caption[]{
\label{fig:carpool}
A \carpool example: (\subref{carpool:input}) an instance containing a
directed graph with capacities on the vertices and weights on the
arcs.  (\subref{carpool:output}) a feasible matching with total weight
of 23.  $P_M$ is the set of blue vertices, and $D_M$ is the set of
red, dashed vertices, and $Z_M$ contains only the dotted, black
vertex.  }
\end{figure}

In this paper we also consider an extension of \carpool, called \gcp,
in which each vertex represents a group of passengers, and each group
might have a different size.  Such a group may be a family or two
friends traveling together.  Formally, each vertex $u \in V$ has a
size $s(u) \in \N$, and the constraint $\din[M](v) \leq c(v)$ is
replaced with the constraint $\sum_{u:(u,v) \in M} s(u) \leq c(v)$.
%
Notice that \textsc{Knapsack} is the special case where only arcs
directed at a single vertex have non-zero (integral) weights.

@@

\subparagraph{Our contribution.}
%
In Sections~\ref{sec:approx} contains approximation algorithms
for \carpool.  First, in Section~\ref{sec:sub} we show that
the \carpool can be formulated as unconstrained submodular
maximization problem, thus it has a $\frac{1}{2}$-approximation
algorithm due to~\cite{BFNS15,buchbinder2016deterministic}.
%
We show a local improvement algorithm for \carpool%
\footnote{We use the term \emph{local improvement} to differentiate
this algorithm from the local search algorithm that is given in
Section~\ref{sec:local}.}.
%
This algorithm repeatedly checks whether the current carpool matching
can be improved by means of a star centered at a vertex, and it
terminates when such a step is not possible.
%
The approximation ratio of this algorithm is $\half$ if weights are
polynomially bounded, and its ratio is $\half-\eps$ in general.

Section~\ref{sec:cmax} we consider \carpool with bounded maximum
capacity.   Define $\cmax \eqdf \max_{v \in V} c(v)$.
%
In Section~\ref{sec:hardness} we show that \carpool is APX-hard even
for undirected and unweighted instances with $\cmax \leq b$, for any
$b \geq 3$.
%
In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide a local search algorithm whose approximation
ratio is $\half + \inv{2\cmax} - \eps$, for any $\eps>0$.  Given a
parameter $K$, our algorithm starts with the empty carpool matching.
Then, it repeatedly tries to find a better matching by replacing
$t \leq k$ arc in the current solution by $t+1$ arcs that are not in
the solution.
%
We show that our analysis is tight.  We also note that our algorithm
is a special case of the local search family that was defined
in~\cite{ACKK09} with a parameter of $2k$.  On undirected and
uncapaciated instances we have that $\cmax = \Delta$, where $\Delta$
is the maximum degree, hence the approximation ratio of our algorithm
is $\half + \inv{2\Delta} - \eps$ which is tight with respect
to~\cite{ACKK09}.

Finally, Section~\ref{sec:group} discusses the \gcp.  We first show that
the unconstrained submodular maximization formulation for the \carpool
does not work for \gcp.  We show, however, that this problem still
admits a $(\frac{1}{2} -\varepsilon)$-approximation algorithm by
extending the local improvement algorithm.  In addition, we show that
the local search algorithm generalizes to \gcp with the same
approximation ratio.

>>>>

\begin{figure}
\centering
\input{fig-carpool}
\caption[]{
\label{fig:carpool}
A \carpool example.
%(\subref{carpool:input}) an instance containing a
%directed graph with capacities on the vertices and weights on the
%arcs.  (\subref{carpool:output}) a feasible matching with total weight
%of 23.  $P_M$ is the set of blue vertices, and $D_M$ is the set of
%red, dashed vertices, and $Z_M$ contains only the dotted, black
%vertex.
}
\end{figure}

%In this paper
%
We also consider an extension of \carpool, called \gcp,
in which each vertex represents a group of passengers, and each group
might have a different size.  Such a group may be a family or two
friends traveling together.  Formally, each vertex $u \in V$ has a
size $s(u) \in \N$, and the constraint $\din[M](v) \leq c(v)$ is
replaced with the constraint $\sum_{u:(u,v) \in M} s(u) \leq c(v)$.
%
Notice that \textsc{Knapsack} is the special case where only arcs
directed at a single vertex have non-zero (integral) weights.

@@

\subparagraph{Our contribution.}
%
Sections~\ref{sec:approx} contains approximation algorithms
for \carpool.  First, in Section~\ref{sec:sub} we show that
the \carpool can be formulated as unconstrained submodular
maximization problem, thus it has a $\frac{1}{2}$-approximation
algorithm due to~\cite{BFNS15,buchbinder2016deterministic}.
%
We present a local search algorithm for \carpool which repeatedly
checks whether the current carpool matching can be improved by means
of a star centered at a vertex, and it terminates when such a step is
not possible.
%
The approximation ratio of this algorithm is $\half$ if weights are
polynomially bounded, and its ratio is $\half-\eps$ in general.

Section~\ref{sec:cmax} we consider \carpool with bounded maximum
capacity.   Define $\cmax \eqdf \max_{v \in V} c(v)$.
%
In Section~\ref{sec:hardness} we show that \carpool is APX-hard even
for undirected and unweighted instances with $\cmax \leq b$, for any
$b \geq 3$.
%
In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide another local search algorithm whose
approximation ratio is $\half + \inv{2\cmax} - \eps$, for any
$\eps>0$.  Given a parameter $k$, our algorithm starts with the empty
carpool matching.  Then, it repeatedly tries to find a better matching
by replacing $t \leq k$ arc in the current solution by $t+1$ arcs that
are not in the solution.
%
We show that our analysis is tight.  We also note that our algorithm
is a special case of the local search family that was defined
in~\cite{ACKK09} with a parameter of $2k$.  On undirected and
uncapaciated instances we have that $\cmax = \Delta$, where $\Delta$
is the maximum degree, hence the approximation ratio of our algorithm
is $\half + \inv{2\Delta} - \eps$ which is tight with respect
to~\cite{ACKK09}.

Finally, Section~\ref{sec:group} discusses the \gcp.  We first show
that the unconstrained submodular maximization formulation for
the \carpool does not work for \gcp.  We show, however, that this
problem still admits a $(\frac{1}{2} -\varepsilon)$-approximation
algorithm by extending our first the local search algorithm.  In
addition, we show that the second local search algorithm generalizes
to unweighted \gcp with the same approximation ratio.
